Applications of the cross-entropy method in diffusion processes

### 2013-12-24 writedown/first.hitting/cross_entropy/cross_entropy.tex
small revise

\begin{table}[t]
  \centering
  \caption{
    Butane dissolved in water. The variance reduced estimate of the probability
    $\mathcal{P} (\tau \leq \lambda)$ by applying control on the dihedral angle.
    See the text for more details of the definition of the control.
    Column ``Error'' denotes the standard deviation of
    estimating the probablity $\mathcal{P} (\tau \leq \lambda)$.
    Column ``Var'' denotes the reduced variation of the esimate.
    ``Accel.''(acceleration) is the ratio between
    the reduced variance and the brute force variance (see Tab.~\ref{tab:tmp2}),
    which is the actual speed-up of the simulation.
    ``Traj. Usage'' dentoes percentage of trajectoies that changes
    to the trans conformation within time interval $[0,\lambda]$.
  }
  \label{tab:tmp1}
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}lcccrr}
    \hline\hline
    $\lambda$ [ps] & $\mathcal{P} (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
    0.1 & $4.34\times 10^{-5}$ & $1.16\times 10^{-5}$ & $2.31\times10^{-6}$ &18.8 & 0.7\%\\
    0.2 & $9.03\times 10^{-4}$ & $0.77\times 10^{-4}$ & $1.19\times10^{-4}$ & 7.6 &10.0\%\\
    0.5 & $6.96\times 10^{-3}$ & $0.34\times 10^{-3}$ & $2.21\times10^{-3}$ & 3.1 &11.9\%\\
    1.0 & $1.68\times 10^{-2}$ & $0.06\times 10^{-2}$ & $6.30\times10^{-3}$ & 2.7 &13.4\%\\
    \hline\hline
  \end{tabular*}
  \caption{
    Butane dissolved in water, brute force simulations
    without any variance reduction. The meaning of the columns are the same
    as Tab.~\ref{tab:tmp1}.}\label{tab:tmp2}
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}lcccrr}
    \hline\hline
    $\lambda$ [ps] & $\mathcal{P} (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
    0.2 & $7.46\times 10^{-4}$ & $2.36\times 10^{-4}$ & $7.46\times10^{-4}$ & 1.0 & 0.1\%\\
    0.5 & $6.94\times 10^{-3}$ & $0.72\times 10^{-3}$ & $6.89\times10^{-3}$ & 1.0 & 0.7\%\\
    1.0 & $1.61\times 10^{-2}$ & $0.11\times 10^{-2}$ & $1.59\times10^{-2}$ & 1.0 & 1.6\%\\
    \hline\hline
  \end{tabular*}
\end{table}

The Fig.~\ref{fig:tmp1} plots the effective dihedral angle energy,
which is the original dihedral energy $V_\phi(\phi)$ plus the control
$V_{\textrm{ctrl}}(\phi)$. We only show the effective energy in the
range $[40^\circ, 150^\circ]$, because the initial
conformations start in range $[40^\circ, 80^\circ]$, and the
trajectoies are stopped when they reach $\phi = 150^\circ$. For an easy
comparison, all effective energies are switched by a constant, so that
they are of vanishing value at $\phi = 150^\circ$. It clear that for
smaller $\lambda$ value, the control applied is stronger. Different
from the one-dimensional case, where the barrier is lowered, in this
case, only the potential well around $\phi = 60^\circ$ is fill, and
the barrier around $\phi = 120^\circ$ is almost kept the same.  The
resultant probability $\mathcal{P} (\tau \leq \lambda)$ calculated by the
variance reduced simulation is summarized in Tab.~\ref{tab:tmp1},
which is consistent with Tab.~\ref{tab:tmp2} that presents the
brute force results. Since the variance is redueced, the error of the
statistical esitmate is more precise in the variance reduced case. The
reduced variance is shown in the fourth column of Tab.~\ref{tab:tmp1},
which is much smaller than the brute force case (the fourth column of
Tab.~\ref{tab:tmp2}). The fifth column presents the acceleration ratio
with respect to the brute force simulation or the benefit from the
variance reduction, which is exactly the ratio between the brute force
variance and the reduced variance. The column of trajectory usage
presents the percentage of the trajectories that changes to the trans
conformation within time interval $[0, \lambda]$. It is increased by
the control with respect to the brute force result.

>>>>

\begin{table}[t]
  \centering
  \caption{
    Butane dissolved in water. The probability
    $\mathcal{P} (\tau \leq \lambda)$ calculated by the important sampling with
    control on the dihedral angle.
    See the text for more details of the definition of the control.
    Column ``Error'' denotes the statistical uncertainty of
    estimating the probablity $\mathcal{P} (\tau \leq \lambda)$.
    If the  trajectories are statistically independent, the expected error is $\sqrt{\textrm{Var}/M}$, where
    $M$ is the number of trajectories. If the trajectories are not independent, the error
    can be estimated by the block average method~\cite{frenkel2001understanding}.
    Column ``Var'' denotes the reduced variance of the esimate.
    ``Accel.''(acceleration) is the ratio between
    the reduced variance and the brute force variance (see Tab.~\ref{tab:tmp2}),
    which is the actual speed-up of the simulation.
    ``Traj. Usage'' dentoes percentage of trajectoies that changes
    to the trans conformation within time interval $[0,\lambda]$.
  }
  \label{tab:tmp1}
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}lcccrr}
    \hline\hline
    $\lambda$ [ps] & $\mathcal{P} (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
    0.1 & $4.34\times 10^{-5}$ & $1.16\times 10^{-5}$ & $2.31\times10^{-6}$ &18.8 & 0.7\%\\
    0.2 & $9.03\times 10^{-4}$ & $0.77\times 10^{-4}$ & $1.19\times10^{-4}$ & 7.6 &10.0\%\\
    0.5 & $6.96\times 10^{-3}$ & $0.34\times 10^{-3}$ & $2.21\times10^{-3}$ & 3.1 &11.9\%\\
    1.0 & $1.68\times 10^{-2}$ & $0.06\times 10^{-2}$ & $6.30\times10^{-3}$ & 2.7 &13.4\%\\
    \hline\hline
  \end{tabular*}
  \caption{
    Butane dissolved in water, brute force simulations
    without any important sampling, i.e.~standard Monte Carlo. The meaning of the columns are the same
    as Tab.~\ref{tab:tmp1}.}\label{tab:tmp2}
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}lcccrr}
    \hline\hline
    $\lambda$ [ps] & $\mathcal{P} (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
    0.2 & $7.46\times 10^{-4}$ & $2.36\times 10^{-4}$ & $7.46\times10^{-4}$ & 1.0 & 0.1\%\\
    0.5 & $6.94\times 10^{-3}$ & $0.72\times 10^{-3}$ & $6.89\times10^{-3}$ & 1.0 & 0.7\%\\
    1.0 & $1.61\times 10^{-2}$ & $0.11\times 10^{-2}$ & $1.59\times10^{-2}$ & 1.0 & 1.6\%\\
    \hline\hline
  \end{tabular*}
\end{table}

The Fig.~\ref{fig:tmp1} plots the effective dihedral angle energy,
which is the original dihedral energy $V_\phi(\phi)$ plus the control
$V_{\textrm{ctrl}}(\phi)$. We only show the effective energy in the
range $[40^\circ, 150^\circ]$, because the initial
conformations start in range $[40^\circ, 80^\circ]$, and the
trajectoies are stopped when they reach $\phi = 150^\circ$. For an easy
comparison, all effective energies are switched by a constant, so that
they are of vanishing value at $\phi = 150^\circ$. It clear that for
smaller $\lambda$ value, the control applied is stronger. Different
from the one-dimensional case, where the barrier is lowered, in this
case, only the potential well around $\phi = 60^\circ$ is fill, and
the barrier around $\phi = 120^\circ$ is almost kept the same.  The
resultant probability $\mathcal{P} (\tau \leq \lambda)$ calculated by the
important sampling simulation is summarized in Tab.~\ref{tab:tmp1},
which is consistent with Tab.~\ref{tab:tmp2} that presents the
brute force results. Since the variance is redueced, the error of the
statistical esitmate is more precise in the important sampling case. The
reduced variance is shown in the fourth column of Tab.~\ref{tab:tmp1},
which is much smaller than the brute force case (the fourth column of
Tab.~\ref{tab:tmp2}). The fifth column presents the acceleration ratio
with respect to the brute force simulation or the benefit from the
important sampling, which is exactly the ratio between the brute force
variance and the reduced variance. The column of trajectory usage
presents the percentage of the trajectories that changes to the trans
conformation within time interval $[0, \lambda]$. It is increased by
the control with respect to the brute force result.
